---
title: "exp011b_analysis"
output: html_notebook
---

based on exp011c load and wrangle

imports isotope corrected excel workbooks, wrangles data, saves tsv output that is then used by `nonlinear fitting` and `plotting` scripts

-import data
-calculate simple fraction (not corrected for natural isotope abundance)
-label samples according to a reference sheet
-plot labeling of each compound

# libraries
```{r}
library(tidyverse)

```


# directories
```{r}

# Set the directory path for free metabolite data
input_free <- "./input/NA corrected/free/"

# output
output_directory <- "./output/natural abundance corrected/"

```



# import data

```{r}



# 1) save free metabolite data to hilic.df



# Get a list of all files in the directory
file_list <- list.files(path = input_free, pattern = "\\.xlsx", full.names=T, recursive=T)

# read each file into a list
data_list <- lapply(file_list, function(x){readxl::read_excel(path=x, sheet="cor_abs")})

# merge all data into a single df
hilic.df <- map_dfr(data_list, identity, .id = "id")




# # 2) save lipidomics data to hilic.df (this is separate because the data is saved in a different folder and the sample names are different)
# source("../../Data Analysis Tools/r scripts/get and merge input csv.R")
# lipids.df <- data.df
# 
# # 3) save hydrolyzed data to hydro.df
# source("../../Data Analysis Tools/r scripts/get and merge input csv.R")
# hydro.df <- data.df

```



# update hilic.df and lipids.df separetely:  melt to tidy format, trim away extra columns

```{r}

hilic.df <-
  hilic.df %>%
    pivot_longer(names_to = "Sample", cols=starts_with("exp0"), values_to = "IonCount") %>%
      select(id, formula, compound, isotopeLabel, Sample, IonCount)

# lipids.df <-
#   lipids.df %>%
#     pivot_longer(names_to = "Sample", cols=starts_with("exp0"), values_to = "IonCount") %>%
#       select(formula, compound, isotopeLabel, Sample, IonCount)
# 
# hydro.df <-
#   hydro.df %>%
#     pivot_longer(names_to = "Sample", cols=starts_with("exp0"), values_to = "IonCount") %>%
#       select(formula, compound, isotopeLabel, Sample, IonCount)
```

# remove duplicates
sometimes I pull the same compound from multiple methods for the same sample.  These duplicates have different values for ion counts.  Here we remove duplicates by assuming the best signal is the observed compound with the best signal.
```{r}

# remove duplicates, save to hilic.df
hilic.df <-
  hilic.df %>% group_by(Sample, compound, id) %>% # id indicates the source file
    summarise(sumIC = sum(IonCount, na.rm=T)) %>%
      filter(sumIC > 0) %>%
      group_by(Sample, compound) %>%
        mutate(best.source = if_else(max(sumIC) == sumIC, id, "0")) %>%
          filter(best.source != "0") %>%
            select(Sample, compound, best.source) %>%
              rename(id = best.source) %>%
    left_join( . , hilic.df)
 



```

Some duplicates are perfect duplicates - this is usually caused by the same compound being pulled twice from the same mzXML (or some glitch in El-Maven that duplicates rows).  I think this is also caused by the same compound being picked from the same mzXML but reported in separate files (ie i forgot i already picked valine before, and i picked it again in a new file).  Here, keep one of the duplicates.
```{r}

# find these duplicates
hilic.df %>% filter(isotopeLabel == "C12 PARENT") %>%
  group_by(Sample, compound, isotopeLabel, IonCount) %>%
    summarise(obs = n()) %>%
      arrange(desc(obs))

# use "distinct" to remove all perfect duplicates.  First i remove the ID column as well.
hilic.df <- hilic.df %>% select(-id) %>% distinct()

# check for duplicates
hilic.df %>% filter(isotopeLabel == "C12 PARENT") %>%
  group_by(Sample, compound, isotopeLabel, IonCount) %>%
    summarise(obs = n()) %>%
      arrange(desc(obs))



```



# update hilic.df and lipids.df:  calculate PeakData fraction for each compound in each sample

```{r}

hilic.df <-
  hilic.df %>%
    group_by(compound, Sample) %>%
      mutate(sumIC = sum(IonCount, na.rm=T),
             fraction.all = IonCount/sumIC) %>%
        ungroup()

# lipids.df <-
#   lipids.df %>%
#     group_by(compound, Sample) %>%
#       mutate(sumIC = sum(IonCount),
#              fraction.all = IonCount/sumIC) %>%
#         ungroup()
# 
# hydro.df <-
#   hydro.df %>%
#     group_by(compound, Sample) %>%
#       mutate(sumIC = sum(IonCount),
#              fraction.all = IonCount/sumIC) %>%
#         ungroup()

```


# import and label sample info for each, then merge into single df
```{r}
sample.reference <- readxl::read_xlsx("../exp011b3_samples.xlsx")

hilic.labeled <- full_join(hilic.df, sample.reference)
# lipids.labeled <- full_join(lipids.df, sample.reference)
# hydro.labeled <- full_join(hydro.df, sample.reference)

# labeled.data <- full_join(hilic.labeled, lipids.labeled)

# labeled.data <- full_join(labeled.data, hydro.labeled)


labeled.data <- hilic.labeled

# filter to remove blanks / samples that were not included ins sample sheet
labeled.data <- labeled.data %>% filter( ! is.na(Method)) %>% filter(! is.na(IonCount))

```




# summarise labeled.data before saving
```{r}

labeled.data %>% summary()


```


# save labeled.data to tsv file used by other tools
```{r}
write_tsv(x = labeled.data, file= paste0(output_directory, "labeled data.tsv"))

# subset of only amino acids from 18O samples, used by Wuhr lab
amino.acids <- c("arginine", "histidine", "lysine", "aspartate", "glutamate", "serine", "threonine", "asparagine", "glutamine", "cysteine", "selenocysteine", "glycine", "proline", "alanine", "valine", "isoleucine", "leucine", "methionine", "phenylalanine", "tyrosine", "tryptophan")

labeled.data %>% filter(compound %in% amino.acids) %>% arrange(Sample, compound, isotopeLabel) %>% select(Sample, compound, isotopeLabel, fraction.all, Tracer, Hours, Tissue) %>%
  write_csv(x= . , file="exp011b3_results_amino acids.csv")

```













